Hashlattice: A distributed system for monotonic peer-to-peer synchronization

### 2022-11-22 paper/hl_paper.tex
graph agda
refactor partial order!

      $$c_1 \leq_{k, v} c_2 \Leftrightarrow \forall k' \in d(k, c_1) \cup d(k, c_2), c_1(k') \leq c_2(k')$$.

@@

    \subsection{Composing semilattice fibrations}

      Semilattice fibrations can be composed iteratively. Composing finitely is straightforward, while omposing infinitely requires more work.

      Suppose we have:

      \begin{align*}
        L_0 &: \mathrm{Semilattice} & \\
        (l_n, t_n) &: \mathrm{SemiLatFib}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
        L_n &:= \mathrm{SemiLatTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
      \end{align*}

      We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$. This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : L_{n-1}$. The joins are defined as follows:

      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)$$

      \begin{theorem}
        $L_{\omega}$ is a semilattice.
      \end{theorem}
      \begin{proof}

        To verify that $L_\omega$ is a semilattice, we must verify three properties:
        \begin{itemize}
          \item \emph{Commutativity}:
            \begin{align*}
              (\overline{x} \vee \overline{y})_n
              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n) \\
              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n})(x_n) \\
              &= (\overline{y} \vee \overline{x})_n
            \end{align*}
          \item \emph{Associativity}:
            \begin{align*}
              & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
                          t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
                          t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)) \\
               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(x_{<n}, x_{<n} \vee y_{<n}))(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, x_{<n} \vee y_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, y_{<n} \vee z_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(z_{<n}, y_{<n} \vee z_{<n}))(z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
                          t_n(y_{<n}, y_{<n} \vee z_{<n})(y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
                          t_n(z_{<n}, y_{<n} \vee z_{<n})(z_n)) \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{y} \vee \overline{z})_n) \\
              =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
            \end{align*}
          \item \emph{Idempotence}:
            \begin{align*}
              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \\
              &= t_n(x_{<n}, x_{<n})(x_n) \\
              &= x_n
            \end{align*}
        \end{itemize}
      \end{proof}

      We can further stack to higher ordinals than $\omega$. Let $n_u \geq 1$ be the number of universes. Assume we have:

      \begin{align*}
        L_{0, 0} &: \mathrm{Semilattice} & \\
        (l_{u, n}, t_{u, n}) &: \mathrm{SemiLatFib}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
        L_{u, n} &:= \mathrm{SemiLatTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
        L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
      \end{align*}

      where $L_{u, \omega}$ is constructed the same way as $L_{\omega}$ above, as the limit of $L_{u, n}$ as $n$ approaches $\infty$. Inductively, each $L_{u, n}$ is a semilattice for naturals $0 \leq u < n_u$ and $n \geq 0$; this lattice's ordinal corresponds to $u \omega + n$. While in theory it may be possible to reach higher ordinals $\omega^2$ and beyond, we consider this beyond the scope of this work.

>>>>

      $$c_1 \leq_{k, v} c_2 \Leftrightarrow \forall k' \in d(k, c_1) \cup d(k, c_2), c_1(k') \leq c_2(k').$$

@@

      A context $c : (K_{<k} \to \mathrm{Option}(V))$ is called \emph{finite} if it is $\mathrm{None}$ for all but a finite set of keys. For $k : K$ we define a join over finite contexts (not yet proven to be a semilattice join):

      $$ (c_1 \vee_k c_2)(k') := c_1(k') \vee_{k', c_1 \vee c_2} c_2(k').$$

      The recursion is well-founded because of the ordering on keys.
